Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-04-04 PandaOverview.tex
Clean up

Currently, PanDA's architecture has five main subsystems: PanDA
Server~\cite{maeno2011overview},
AutoPyFactory~\cite{caballero2012autopyfactory}, PanDA
Pilot~\cite{nilsson2011atlas}, JEDI~\cite{borodin2015scaling}, and PanDA
Monitoring~\cite{klimentov2011atlas}. Other subsystems are used by some of
ATLAS workflows (e.g., PanDA Event Service~\cite{calafiura2015atlas}), but we
do not discuss them as they are  not relevant to  an understanding of how
PanDA has been ported to supercomputers. For a full list of subsystems see
Ref.~\cite{panda-wiki_url}. Figure~\ref{fig:architecture} shows a diagrammatic
representation of PanDA main subsystems, highlighting the execution process of
tasks while omitting monitoring details to improve readability.

>>>>

Currently, PanDA's architecture has five main subsystems: PanDA
Server~\cite{maeno2011overview},
AutoPyFactory~\cite{caballero2012autopyfactory}, PanDA
Pilot~\cite{nilsson2011atlas}, JEDI~\cite{borodin2015scaling}, and PanDA
Monitoring~\cite{klimentov2011atlas}. Other subsystems are used by some of ATLAS
workflows (e.g., PanDA Event Service~\cite{calafiura2015atlas}), but we do not
discuss them as they are  not relevant to  an understanding of how PanDA has
been ported to supercomputers. For a full list of subsystems see
Ref.~\cite{panda-wiki_url}. Fig.~\ref{fig:architecture} shows a diagrammatic
representation of PanDA main subsystems, highlighting the execution process of
tasks while omitting monitoring details to improve readability.

### 2017-04-04 nge.tex
Clean up

% \subsection{Rationale}
% \aanote{I think that the names of the states should be removed. They are
% meaningless without the figures. I left them just in case we decide to
% reintroduce the figures.}\mtnote{Done.}

% \jhanote{admittedly i haven't read earlier section in question  carefully, but
% i'm a bit confused about the use of "wide area". wide area is typically used for
% cross data center}\mtnote{Every PanDA Pilot has to communicate with PanDA
% Server. PanDA Server runs on dedicated resources, outside OLCF. Titan's worker
% nodes cannot communicate with PanDA Server because their connectivity is limited
% to Titan's network. Do you have a suggestion for an alternative to ``wide
% area''?} \jhanote{how is the following: "... deployed on the DTNs as Titan's worker
% nodes lack connectivity to the wide area network"}.

@@

% \mtnote{Are there more reasons for wanting multiple generations?}

% \jhanote{not sure MPI script is optimal description or ambiguous.}\mtnote{Is
% what we used in the previous section(s) to designate what PanDA Broker submits
% to Titan's PBS batch system. Job is already used for what is pulled from
% PanDA server.}

Multiple generations would also relax two assumptions of the current execution
model: knowing the number of simulations before submitting the MPI script, and
having a fixed number of events per simulation (100 at the moment). Pilots would
enable the scheduling of simulations independently from whether they were
available at the moment of submitting the pilot. Further, simulations with a
varying number of events could be scheduled on a pilot, depending on the amount
of remaining walltime and the distribution of execution time per event, as shown
in \S\ref{ssec:panda_titan}, Figure~\ref{fig:100event- distrib}. These
capabilities would increase the efficiency of the PanDA Broker when there is a
large difference between the number of cores and walltime.

% to one in which the number of events is decided on the base of the available
% walltime. Pilots would enable to package simulations with a heterogeneous
% number of events, especially when a small amount of cores is available for a
% long period of time and \textit{vice versa}. In these situations, a PBS job
% would require all simulations to have a small number of events that would not
% use all the available walltime.

% a pilot would instead support multiple generations of simulations with a
% different number of events and therefore duration.

@@

% \mtnote{I tried to argue that pilots facilitates the submission to a normal
% queue but I was not able to find a convincing argumentation. When we consider
% only Titan, a PBS job is the only way to submit a job to its queue(s). With %
% NGE, we translate a pilot description to a PBS job and then submit it to a
% queue via a PBS command. PanDA Brokers do something very similar: they
% package detector simulations into a PBS job and they submit it via a PBS
% command. The fact that PanDA Brokers define the number of cores and walltime
% of a PBS job on the base of backfill information, does not make the packaging
% or the submission processes any different: PanDA Brokers are already
% submitting to the nomal queue. If I got it right, PanDA Brokers could already
% submit a job with, say, 12000 cores and 2 hours walltime. They don't do it
% because they have no allocation and they would wait a long time in the queue
% (maybe, let's see what our experiments say about that). I do not see how NGE
% could make all this any different.}

% \mtnote{NOTE: I would speak about about backfill/not backfill in the discussion
% of NGE, when speaking about its generality towards resources, i.e., unified
% submission and scheduling process across different resources and multiple
% resources. Part of this generality is being able to submit to whatever batch
% system is supported by SAGA (including Titan's PBS and all its queues) and, in
% case, to multiple resources at the same time.}
%
% \mtnote{NOTE: NGE is not able to use backfill on Titan as we do not have
% specific functionalities to interrogate the Moab scheduler about backfill
% availability. This is why NGE should be presented as the pilot for PanDA Broker
% and not as an alternative to PanDA Broker. Further, NGE alone would not be able
% to speak directly to PanDA Server.}

@@

% We developed and we are currently integrating a prototype of Next Generation
% Executor (NGE) to add pilot capabilities to the PanDA Broker.

@@

% Thus, workloads and pilots are described via the Pilot API and passed to the
% NGE runtime system. NGE launches the pilots and executes the tasks of the
% workload on them.

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{figures/rp_architecture_compact_atlaswms_paper.pdf}
  % \caption{RP overview. An application uses the Pilot API to describe pilots
  %   (green squares) and units (red circles). The PilotManager (purple box)
  %   instantiates (dash arrow) pilots, the UnitManager (purple box)
  %   instantiates (solid arrow) units. Both managers are executed on the user
  %   workstation. Pilots are launched (dashed arrows) on resources A and B
  %   via SAGA API, an Agent (purple box) is bootstrapped for each pilot
  %   (green box), units (solid red circles) are scheduled to the Agents via
  %   MongoDB and executed by the Agent's Executer (yellow box).
  \caption{NGE Architecture as deployed on Titan. The PilotManager and the
  UnitManager reside on one of Titan's DTN while the Agent is executed on
  Titan's worker nodes. Boxes color coding: gray for entities external to NGE,
  white for APIs, purple for NGE's modules, green for pilots, yellow for
  module's components.}
\label{fig:arch-overview}
\end{figure}

% Internally, NGE represents pilots as aggregates of resources independent from
% the architecture and topology of the target machines, and workloads as a set
% of units to be executed on the resources of the pilot. Both pilots and units
% are stateful entities, each with a well-defined state model and life cycle.
% Their states and state transitions are managed via the three modules of the
% NGE architecture: PilotManager, UnitManager, and Agent
% (Fig.~\ref{fig:arch-overview}, purple boxes).

@@

% \jhanote{for consistency should we say Moab?}\mtnote{Moab = Scheduler; Torque
% = Batch system. I used PBS Batch System because it makes easier to speak about
% both the system and the script submitted to it.}

% A similar approach would be used to enable PanDA Broker to schedule pilots
% instead of MPI scripts.\mtnote{Rewrite this after checking it with
% Andre.}
% \jhanote{In the interest of space and clarity, I would  remove the last
% sentence of this para. And merge the first sentence with the para
% above.}\mtnote{Done.}

% The lifespan of pilots has four states distributed among the PilotManager,
% resource, and pilot instance.% (Fig.~\ref{fig:pilot-state-model}).
% Pilots are instantiated in the state \T{NEW} by the PilotManager, wait in a
% queue to be launched, and transition to \T{PM\_LAUNCH} when submitted to a
% Resource Manager (RM) via the SAGA API\@. Pilots wait in the queue of the RM
% and, once scheduled, become \T{P\_ACTIVE}. They remain in this state until the
% end of their lifetime, when they transition to \T{DONE}.

%\begin{figure}
%  \centering
%  \includegraphics[width=\columnwidth]{figures/pilot_state_model_compact_rp_paper.pdf}
%  \caption{Pilot State Model. Instantiated in state \T{NEW}, each pilot is
%    launched (\T{PM\_LAUNCH}) via SAGA API on a resource manager. After
%    becoming \T{P\_ACTIVE}, it exhausts its duration ending in state
%    \T{DONE}. The transition between states is sequential and each transition
%    can be canceled or fail ending in the states \T{CANCELED} or \T{FAILED} (not
%    depicted to improve the diagram clarity). Boxes color coding as per
%    Fig.~\ref{fig:arch-overview}, with: orange for queues and blue for states.
%\label{fig:pilot-state-model}}
%\end{figure}
%

% The unit state model has nine states distributed across the UnitManager,
% MongoDB instance, and Agent.% (Fig.~\ref{fig:unit-state-model}).
% Instantiated in the state \T{NEW} by the UnitManager, every unit is scheduled
% on an Agent (\T{UM\_SCHEDULING}) via a queue on a MongoDB instance. The unit
% is then scheduled on the required number of cores held by the Agent's pilot
% (\T{A\_SCHEDULING}), and finally executed (\T{A\_EXECUTING}).

% The unit state model pertains also to the input and output data of the units.
% When required, the input data of a unit are either pushed to the Agent
% (\T{U\_STAGING\_IN}) or pulled from the Agent (\T{A\_STAGING\_IN}), depending
% on data locality and sharing requirements. Similarly, the output data of the
% unit are staged out by the Agent and UnitManager (\T{A\_STAGING\_OUT},
% \T{U\_STAGING\_OUT}) to a specified destination, e.g., the user workstation.
% Both input and output staging are optional, depending on the requirements of
% the units. The actual file transfers are enacted via SAGA, and support
% (gsi)-scp, (gsi)-sftp, and Globus Online.

%\begin{figure}
%  \centering
%  \includegraphics[width=\columnwidth]{figures/unit_state_model_compact_rp_paper.pdf}
%  \caption{Unit State Model. Instantiated in state \T{NEW}, each unit is
%    scheduled to an Agent (\T{UM\_SCHEDULING}) via MongoDB, and then scheduled
%    to an Agent's Executer (\T{A\_SCHEDULING}). When required, unit input data
%    are staged to the Agent's Pilot (\T{UM\_STAGING\_IN} or
%    \T{A\_STAGING\_IN}). Each unit is executed (\T{A\_EXECUTING}), unit output
%    data are staged out (\T{A\_STAGING\_OUT} and \T{UM\_STAGING\_OUT}), and
%    each unit ends in state (\T{DONE}). Color coding and state omissions
%    as per Fig.~\ref{fig:pilot-state-model}.
%\label{fig:unit-state-model}}
%\end{figure}

%The state transitions represented in Figures~\ref{fig:pilot-state-model}
%and~\ref{fig:unit-state-model} are sequential and every transition can fail or
%be canceled by the PilotManager or UnitManager.

% All state transitions are managed by the PilotManager, UnitManager, and Agent
% components. The only special case is the transition of the pilots to the state
% \T{P\_ACTIVE} which is dictated by the resource's RM, but managed by the
% PilotManager.

% Above all NGE's components, the Agent is the one that is worth to mention the
% most.

@@

% ORTE transparently provides support for interprocess communication, resource
% discovery and allocation, and process launching across a variety of platforms.
% ORTE provides a mechanism similar to the Pilot concept - it allows the user to
% create a \emph{``dynamic virtual machine''} (DVM) that spans multiple nodes.
% ORTE provides libraries to enable the submission, monitoring and managing of
% tasks, avoiding filesystem bottlenecks and race conditions with network sockets.
% \aanote{Requires validation: Alternative version of the three sentences above
% (too many ORTE in my opinion):}

%Each component of RP has a independent semantic scope. This enables modularity
%isolating implementation complexity and supporting diverse use cases and
%environments.  For example, unit scheduling can be implemented by exchangeable
%Scheduler components, suitable for applications of diverse scales, with
%different coordination patterns, and executed on Beowulf clusters or Cray
%machines.
%
%Components are also designed to be stateless and instantiated concurrently. In
%this way, RP can manage multiple pilots and units at the same time, resulting
%in scalable throughput and tolerance to failing components. Concurrent
%components are coordinated via a dedicated communication mesh which incurs
%infrastructure and runtime overhead, offset by the lower component complexity
%and improved overall scalability of the system.

% \begin{enumerate}
%     \item Rationale and design (why)
%     \item Architecture (how)
%     \item Integration (Not present)
%     \item Characterization (experiments)
% \end{enumerate}

 % \ldots

>>>>

@@

Multiple generations would also relax two assumptions of the current execution
model: knowing the number of simulations before submitting the MPI script, and
having a fixed number of events per simulation (100 at the moment). Pilots would
enable the scheduling of simulations independently from whether they were
available at the moment of submitting the pilot. Further, simulations with a
varying number of events could be scheduled on a pilot, depending on the amount
of remaining walltime and the distribution of execution time per event, as shown
in \S\ref{ssec:panda_titan}, Fig.~\ref{fig:100event- distrib}. These
capabilities would increase the efficiency of the PanDA Broker when there is a
large difference between the number of cores and walltime.

@@

@@

@@

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{figures/rp_architecture_compact_atlaswms_paper.pdf}
  \caption{NGE Architecture as deployed on Titan. The PilotManager and the
  UnitManager reside on one of Titan's DTN while the Agent is executed on
  Titan's worker nodes. Boxes color coding: gray for entities external to NGE,
  white for APIs, purple for NGE's modules, green for pilots, yellow for
  module's components.}
\label{fig:arch-overview}
\end{figure}

@@

@@
